State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2024-05-27 StateOfPractice/Papers/SOP_MI_OSR/SOP_MI_OSR.tex
Split versions of SOP papers

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for Elsevier's document class `elsarticle'
%% with Harvard style bibliographic references

%\documentclass[preprint,12pt,authoryear]{elsarticle}
%\documentclass[3p, 12pt,authoryear]{elsarticle}
\documentclass[final, 3p, times, authoryear, doubleblind]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times,authoryear]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{amsmath,amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

@@

\journal{Archives of Computational Methods in Engineering}

@@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \affiliation for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%            addressline={}, 
%%            city={},
%%            postcode={}, 
%%            state={},
%%            country={}}
%% \fntext[label3]{}

\title{State of the Practice for Medical Imaging Software}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[CAS]{Spencer Smith}
\author[CAS]{Ao Dong}
\author[CAS]{Jacques Carette}
\author[ECE]{Mike Noseworthy}

\affiliation[CAS]{organization={McMaster University, Computing and Software Department},%Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

\affiliation[ECE]{organization={McMaster University, Electrical Engineering},%Department and Organization
            addressline={1280 Main Street West}, 
            city={Hamilton},
            postcode={L8S 4K1}, 
            state={Ontario},
            country={Canada}}

@@

We present the state of the practice for Medical Imaging (MI) software. We
selected 29 medical imaging projects from 48 candidates, assessed 10 software
qualities (installability, correctness/ verifiability, reliability, robustness,
usability, maintainability, reusability, understandability,
visibility/transparency and reproducibility) by answering 108 questions for each
software project, and interviewed 8 of the 29 development teams. Based on the
quantitative data for the first 9 qualities, we ranked the MI software with the
Analytic Hierarchy Process (AHP). The four top ranked software products are:
\textit{3D Slicer}, \textit{ImageJ}, \textit{Fiji}, and \textit{OHIF Viewer}.
Our ranking is mostly consistent with the community's ranking, with four of our
top five projects also appearing in the top five of a list ordered by
stars-per-year. Generally, MI software is in a healthy state as shown by the
following: in the repositories we observed 88\% of the documentation artifacts
recommended by research software development guidelines, 100\% of MI projects
use version control tools, and developers appear to use the common quasi-agile
research software development process. However, the current state of the
practice deviates from the existing guidelines because of the rarity of some
recommended artifacts (like test plans, requirements specification, code of
conduct, code style guidelines, product roadmaps, and Application Program
Interface (API) documentation), low usage of continuous integration (17\% of the
projects), low use of unit testing (about 50\% of projects), and room for
improvement with documentation (six of nine developers felt their documentation
wasn't clear enough). From interviewing the developers, we identified five pain
points and two qualities of potential concern: lack of development time, lack of
funding, technology hurdles, ensuring correctness, usability, maintainability,
and reproducibility. The interviewees proposed strategies to improve the state
of the practice, to address the identified pain points, and to improve software
quality. Combining their ideas with ours, we have the following list of
recommendations: increase documentation, increase testing by enriching datasets,
increase continuous integration usage, move to web applications, employ linters,
use peer reviews, design for change, add assurance cases, and incorporate a
``Generate All Things'' approach.

@@

We aim to study the state of software development practice for Medical Imaging
(MI) software.  MI tools use images of the interior of the body (from sources
such as Magnetic Resonance Imaging (MRI), Computed Tomography (CT), Positron
Emission Tomography (PET) and Ultrasound) to provide information for diagnostic,
analytic, and medical applications \citep{FDA2021, enwiki:1034887445,
Zhang2008}.  Figure~\ref{Fig_Example}, which shows an image of the brain,
highlights the importance and value of MI. Through MI medical practitioners and
researchers can noninvasively gain insights into the human body, including
information on injuries and illnesses. Given the importance of MI software and
the high number of competing software projects, we wish to understand the merits
and drawbacks of the current development processes, tools, and methodologies. We
aim to assess through a software engineering lens the quality of the existing
software with the hope of highlighting standout examples, understanding current
pain points and providing guidelines and recommendations for future development.

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[scale=0.25]{figures/MPR.png}        
    \end{center}
    \caption{Example brain image showing a multi-planar reformat using Horos
	(free open-source medical imaging/DICOM viewer for OSX, based on OsiriX)}
    \label{Fig_Example}
\end{figure}

@@

Domain experts may be recruited from academia or industry.  The only
requirements are knowledge of the domain and a willingness to be engaged in the
assessment process.  The Domain Expert does not have to be a software developer,
but they should be a user of domain software.  Given that the domain experts are
likely to be busy people, the measurement process cannot put too much of a burden
on their time.  For the current assessment, our Domain Expert (and paper
co-author) is Anonymous % DOUBLE BLIND
% Dr.\ Michael Noseworthy, Professor of Electrical and Computer
% Engineering at McMaster University, Co-Director of the McMaster School of
% Biomedical Engineering, and Director of Medical Imaging Physics and Engineering
% at St.\ Joseph's Healthcare, Hamilton, Ontario, Canada.

@@

We based our interviews on a list of 20 questions, which can be found in
\citet{SmithEtAl2021}. Some questions are about the background of the software,
the development teams, the interviewees, and how they organize their projects.
We also ask about the developer's understanding of the users. Some questions
focus on the current and past difficulties, and the solutions the team has
found, or plan to try. We also discuss documentation, both with respect to how
it is currently done, and how it is perceived. A few questions are about
specific software qualities, such as maintainability, understandability,
usability, and reproducibility. The interviews are semi-structured based on the
question list; we ask follow-up questions when necessary.  The interview process
presented here was approved by the Anonymous Ethics Board. %% DOUBLE BLIND
% McMaster University Research Ethics Board
% under the application number 
% \href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
% {MREB\#: 5219}.

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.5]{figures/PrimaryLanguages.pdf}
\centering
\caption{\label{fig_language}Primary languages versus number of projects}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/installability_scores.pdf}
\caption{AHP installability scores}
\label{fg_installability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/correctness_verifiability_scores.pdf}
\caption{AHP correctness \& verifiability scores}
\label{fg_correctness_verifiability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/reliability_scores.pdf}
\caption{AHP surface reliability scores}
\label{fg_reliability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/robustness_scores.pdf}
\caption{AHP surface robustness scores}
\label{fg_robustness_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/usability_scores.pdf}
\caption{AHP surface usability scores}
\label{fg_usability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/maintainability_scores.pdf}
\caption{AHP maintainability scores}
\label{fg_maintainability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/reusability_scores.pdf}
\caption{AHP reusability scores}
\label{fg_reusability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/understandability_scores.pdf}
\caption{AHP surface understandability scores}
\label{fg_surface_understandability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/visibility_transparency_scores.pdf}
\caption{AHP visibility/transparency scores}
\label{fg_visibility_transparency_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{figures/overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \citet{USGS2019} & \citet{TobiasEtAl2018} & \citet{BrettEtAl2021} &
\citet{WilsonEtAl2016} & \citet{SmithAndRoscoe2018} & \citet{HerouxEtAl2008} &
\citet{ThielEtAl2020} & \citet{vanGompelEtAl2016} & \citet{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark &
\checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark & R\\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table}

@@

\section{Comparison of Tool Usage Between MI and Other Research Software}
\label{Sec_CompareTools}

Developers use software tools to support the development, verification,
maintenance, and evolution of software, software processes, and artifacts
\citep[p.\ 501]{GhezziEtAl2003}. MI software uses tools for CI/CD, user support,
version control, documentation, and project management.  To answer
\rqref{RQ_CompareToolsProjMngmnt} we summarize the tool usage in these
categories, and compare this to the usage by the research software community.

Table~\ref{tab_user_support_model} summarizes the user support models by the
number of projects using each model (projects may use more than one support
model). We do not know whether the prevalent use of GitHub issues for user
support is by design, or whether this just naturally happens as users seek
help. The common use of GitHub by MI developers is not surprising, given that
GitHub is the largest code host in the world, with over 128 million public
repositories and over 23 million users (as of roughly February 26, 2020)
\citep{Kashyap2020}.

\begin{table}[!ht]
\centering
\begin{tabular}{lc}
\toprule
\multicolumn{1}{c}{User Support Model} & Num.\ Projects \\
\midrule
GitHub issue & 24 \\
Frequently Asked Questions (FAQ) & 12 \\
Forum & 10 \\
E-mail address & 9 \\
GitLab issue, SourceForge discussions & 2 \\
Troubleshooting & 2 \\
Contact form & 1 \\ 
\bottomrule
\end{tabular}
\caption{\label{tab_user_support_model}User support models by number of projects}
\end{table}

From Section~\ref{sec_score_maintainability}, 27 of the 29 projects used git as
the version control tool, one used Mercurial and one used Subversion.  The
hosting is on GitHub for 24 packages, SourceForge for three and BitBucket for
two.  Although teams may have a process for accepting new contributions, no one
discussed this during their interviews. However, most teams (eight of nine)
mentioned using GitHub and pull requests to manage contributions from the
community. The interviewees generally gave very positive feedback on using
GitHub. Some teams previously used a different approach to version control and
eventually transferred to git and GitHub.  The past approaches included
contributions from e-mail (three teams), contributions from forums (one team)
and e-mailing the git repository back and forth between developers (one team).

The common use of version control for MI software illustrates considerable
improvement from the poor adoption of version control tools that Wilson lamented
in 2006 \citep{Wilson2006}.  The proliferation of version control tools for MI
matches the increase in the broader research software community.  A little over
10 years ago \citet{Nguyen-HoanEtAl2010} estimated that only 50\% of research
software projects use version control, but even at that time
\citet{Nguyen-HoanEtAl2010} noted an increase from previous usage levels. A
survey in 2018 shows 81\% of developers use a version control system
\citep{AlNoamanyAndBorghi2018}. \citet{Smith2018} has similar results, showing
version control usage for alive projects in mesh generation, geographic
information systems and statistical software for psychiatry increasing from
75\%, 89\% and 17\% (respectively) to 100\%, 95\% and 100\% (respectively) over
a four-year period ending in 2018. (For completeness the same study showed a
small decrease in version control usage for seismology software over the same
time period, from 41\% down to 36\%).  A recent survey by \citet{CarverEtAl2022}
shows version control use among practitioners at over 95\%, with 83/87 survey
respondents indicating that they use it. All but one of the software guides
cited in Section~\ref{Sec_CompareArtifacts} includes the advice to use version
control. (The USGS guide \citep{USGS2019} was the only set of recommendations to
not mention version control.) The high usage of version control tools in MI
software matches the trend in research software in general.

As mentioned in Section~\ref{sec_result_correctness_verifiability}, we
identified five projects using CI/CD tools (about 17\% of the assessed
projects). We found which projects used CI/CD by examining the documentation and
source code of all projects. The count of CI/CD usage may actually be higher,
since traces of CI/CD usage may not always appear in a repository.  This was the
case for a study of LBM software, where interviews with developers showed that
more projects used CI/CD than was evident from repository artifacts alone
\citep{Michalski2021}.  The 17\% utilization for MI software contrasts with the
high frequency with which research software development guidelines recommend
continuous integration \citep{BrettEtAl2021, Brown2015, ThielEtAl2020,
Zadka2018, vanGompelEtAl2016}. Although there is currently little data available
on CI/CD utilization for research software, our impression is that CI/CD is not
yet common practice, despite its recommendation.  This is certainly the case for
LBM software, where usage numbers are similar to MI software, with only 12.5\%
of 24 LBM packages showing evidence of CI/CD in their repositories
\citep{Michalski2021}.  The survey of \citet{CarverEtAl2022} suggests higher use
of CI/CD with 54\% (54/100) of respondents indicating that they use it. However,
that survey measures something different from the current one by surveying
practitioners, rather than assessing projects.  Additional information on CI/CD
is given in the recommendations (Section~\ref{Sec_ContinuousIntegration}).

For documentation tools and methods mentioned by the interviewees, the most
popular (mentioned by about 30\% of developers) were forum discussions and
videos.  The second most popular options (mentioned by about 20\% of developers)
were GitHub, wiki pages, workshops, and social media. The least frequently
mentioned options (about 10\% of developers) included writing books, and google
forms.  In contrasting MI software with LBM software, the most significant
documentation tool difference is that LBM software often uses document
generation tools, like doxygen and sphinx \citep{Michalski2021}, while MI does
not appear to use these tools.

Some interviewees mentioned the project management tools they used. Generally
speaking, the interviewees talked about two types of tools:
\begin{inparaenum}[i)]
\item trackers, including GitHub, issue trackers, bug trackers and Jira; and,
\item documentation tools, including GitHub, Wiki page, Google Doc, and
Confluence.
\end{inparaenum}
Of the specifically named tools in the above lists, interviewees mentioned
GitHub 3 times, and each of the other tools once each.

Based on information provided by \citet{JungEtAl2022}, tool utilization for MI
software has much in common with tool utilization for ocean modelling software.
Both use tools for editing, compiling, code management, testing, building, and
project management.  From the data available, ocean modelling differs from MI
software in its use of Kanban boards for project management.

\section{Comparison of Principles, Process, and Methodologies to Research Software in General} \label{Sec_CompareMethodologies}

We answer research question \rqref{RQ_CompareMethodologies} by comparing the
principles, processes, and methodologies used for MI software to what can be
gleaned from the literature on research software in general. In our interviews
with developers the responses about development model were vague, with only two
interviewees following a definite development model. In some cases the
interviewees felt their process was similar to an existing development model.
Three teams (about 38\%) either followed agile, or something similar to agile.
Two teams (25\%) either followed a waterfall process, or something similar.
Three teams (about 38\%) explicitly stated that their process was undefined or
self-directed.

Our observations of an informally defined process, with elements of agile
methods, matches what has been observed for research software in general.
Scientific developers naturally use an agile philosophy \citep{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \citep{Kelly2013}, or a knowledge acquisition driven
process \citep{Kelly2015}.  A waterfall-like process can work for research
software \citep{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a
rationale design process \citep{parnas1986rational}.

No interviewee introduced any strictly defined project management process. The
most common approach was following the issues, such as bugs and feature
requests. Additionally, the \textit{3D Slicer} team had weekly meetings to
discuss the goals for the project; the \textit{INVESALIUS 3} team relied on the
GitHub process for their project management; the \textit{ITK-SNAP} team had a
fixed six-month release pace; only the interviewee from the \textit{OHIF} team
mentioned that the team has a project manager; the \textit{3D Slicer} team and
\textit{BioImage Suite Web} team do nightly builds and tests. The \textit{OHIF}
developer believes that a better project management process can improve junior
developer efficiency while also improving internal and external communication.

We identified the use of unit testing in less than half of the 29 projects. On
the other hand, the interviewees believed that testing (including usability
tests with users) was the top solution to improve correctness, usability, and
reproducibility.  This level of testing matches what was observed for LBM
software \citep{Michalski2021} and is apparently greater than the level of
testing for ocean modelling software.  \citet{JungEtAl2022} reports that ocean
modellers underemphasize testing.

As the observed artifacts in Table~\ref{artifactspresent} show, none of the 29
projects emphasize documentation. None of them had theory manuals, although we
did identify a road map in the \textit{3D Slicer} project.  We did not find
requirements specifications. Table~\ref{tab_opinion_doc} summarizes
interviewees' opinions on documentation. Interviewees from each of the eight
projects thought that documentation was essential to their projects, and most of
them said that it could save their time to answer questions from users and
developers. Most of them saw the need to improve their documentation, and only
three of them thought that their documentations conveyed information clearly
enough. Nearly half of developers also believed that the lack of time prevented
them from improving documentation.

\begin{table}[!ht]
\centering
\begin{tabular}{ll}
\toprule
Opinion on Documentation & Num.\ Ans. \\ 
\midrule
Documentation is vital to the project & 8 \\
Documentation of the project needs improvements & 7 \\
Referring to documentation saves time to answer questions & 6 \\
Lack of time to maintain good documentation & 4 \\
Documentation of the project conveys information clearly & 3 \\
Coding is more fun than documentation & 2 \\
Users help each other by referring to documentation & 1 \\ 
\bottomrule
\end{tabular}
\caption{Opinions on documentation by the numbers of interviewees with the
answers}
\label{tab_opinion_doc}
\end{table}

As Table~\ref{Tbl_Guidelines} suggests, an emphasis on documentation, especially
for new developers, is echoed in research software guidelines. Multiple
guidelines recommend a document explaining how to contribute to a project, often
named CONTRIBUTING. Guidelines also recommend tutorials, user guides and quick
start examples. \citet{SmithAndRoscoe2018} suggests including instructions
specifically for on-boarding new developers. For open-source software in general
(not just research software), \citet{Fogel2005} recommends providing tutorial
style examples, developer guidelines, demos, and screenshots.

\section{Developer Pain Points} \label{painpoints}

Based on interviews with nine developers (described in
Section~\ref{sec_interview_methods}), we answer three research questions (first
mentioned in Section~\ref{sec_motivation}): \rqref{RQ_PainPoints}) What are the
pain points for developers working on research software projects?;
\rqref{RQ_ComparePainPoints}) How do the pain points of developers from MI
compare to the pain points for research software in general?; and
\rqref{RQ_Concerns}) For MI developers what specific best practices are taken to
address the pain points and software quality concerns?

Our interviews identified pain points related to a lack of time and funding,
technology hurdles, improving correctness, and improving usability.  In this
section, we go through each pain point and contrast the MI experience with
observations from other domains.  We also cover potential ways to address the
pain points, as promoted by the community.  (Later, in
Section~\ref{ch_recommendations}, we propose additional pain mitigation
strategies based on our experience.)  In addition to pain points, we summarize
MI developer strategies for improving maintainability and reproducibility.
Although the interviewees did not explicitly identify these two qualities as
pain points, we did discuss threats to these qualities and ways to improve them
as part of our interview process \citep{SmithEtAl2021}.  The interviewee's
practices for addressing pain points and improving quality can potentially be
emulated by other MI developers. Moreover, these practices may provide examples
that can be followed by other research software domains.

>>>>

\documentclass[11pt]{article}

\usepackage{authblk}
\usepackage{fancyhdr}
\usepackage{float}

\usepackage[round, sort]{natbib}

\usepackage{amssymb}
\usepackage{amsmath,amsthm}

@@

@@

\title{State of the Practice for Medical Imaging Software Based on Open Source Repositories}

\author[1,*]{W.\ Spencer Smith}
\author[1]{Ao Dong}
\author[1]{Jacques Carette}
\author[2]{Michael D.\ Noseworthy}

\affil[1]{McMaster University, Computing and Software Department, {Canada}}
\affil[2]{McMaster University, Electrical \& Computer Engineering Department,
{Canada}}
\affil[*]{Corresponding Author}

\maketitle

% RQ1 to RQ4
% Some of Section 1 (Introduction)
% Sections 2 to 6 (includes looking at artifacts observed in repos)
% Some of Sections 11 to 13 (Threats to Validity, Future Work, Conclusions)

@@

We present the state of the practice for Medical Imaging (MI) software based on
data available in open source repositories. We selected 29 medical imaging
projects from 48 candidates and assessed 10 software qualities (installability,
correctness/ verifiability, reliability, robustness, usability, maintainability,
reusability, understandability, visibility/transparency and reproducibility) by
answering 108 questions for each software project. Based on the quantitative
data for the first 9 qualities, we ranked the MI software with the Analytic
Hierarchy Process (AHP). The four top ranked software products are \textit{3D
Slicer}, \textit{ImageJ}, \textit{Fiji}, and \textit{OHIF Viewer}.  Our ranking
is mostly consistent with the community's ranking, with four of our top five
projects also appearing in the top five of a list ordered by stars-per-year.  We
observed 88\% of the documentation artifacts recommended by research software
development guidelines.  However, the current state of the practice deviates
from the existing guidelines because of the rarity of some recommended artifacts
(like test plans, requirements specification, code of conduct, code style
guidelines, product roadmaps, and Application Program Interface (API)
documentation). Open source MI software projects may wish to consider increasing
their documentation.

@@

We aim to study the state of software development practice for Medical Imaging
(MI) software using data available in open source repositories.  MI tools use
images of the interior of the body (from sources such as Magnetic Resonance
Imaging (MRI), Computed Tomography (CT), Positron Emission Tomography (PET) and
Ultrasound) to provide information for diagnostic, analytic, and medical
applications \citep{FDA2021, enwiki:1034887445, Zhang2008}.
Figure~\ref{Fig_Example}, which shows an image of the brain, highlights the
importance and value of MI. Through MI medical practitioners and researchers can
noninvasively gain insights into the human body, including information on
injuries and illnesses. Given the importance of MI software and the high number
of competing software projects, we wish to understand the merits and drawbacks
of the current development processes, tools, and methodologies. We aim to assess
through a software engineering lens the quality of the existing software with
the hope of highlighting standout examples, understanding current pain points
and providing guidelines and recommendations for future development.

\begin{figure}[!ht]
    \begin{center}
        \includegraphics[scale=0.25]{MPR.png}        
    \end{center}
    \caption{Example brain image showing a multi-planar reformat using Horos
	(free open-source medical imaging/DICOM viewer for OSX, based on OsiriX)}
    \label{Fig_Example}
\end{figure}

@@

Domain experts may be recruited from academia or industry.  The only
requirements are knowledge of the domain and a willingness to be engaged in the
assessment process.  The Domain Expert does not have to be a software developer,
but they should be a user of domain software.  Given that the domain experts are
likely to be busy people, the measurement process cannot put too much of a burden
on their time.  For the current assessment, our Domain Expert (and paper
co-author) is Dr.\ Michael Noseworthy, Professor of Electrical and Computer
Engineering at McMaster University, Co-Director of the McMaster School of
Biomedical Engineering, and Director of Medical Imaging Physics and Engineering
at St.\ Joseph's Healthcare, Hamilton, Ontario, Canada.

@@

We based our interviews on a list of 20 questions, which can be found in
\citet{SmithEtAl2021}. Some questions are about the background of the software,
the development teams, the interviewees, and how they organize their projects.
We also ask about the developer's understanding of the users. Some questions
focus on the current and past difficulties, and the solutions the team has
found, or plan to try. We also discuss documentation, both with respect to how
it is currently done, and how it is perceived. A few questions are about
specific software qualities, such as maintainability, understandability,
usability, and reproducibility. The interviews are semi-structured based on the
question list; we ask follow-up questions when necessary.  The interview process
presented here was approved by the McMaster University Research Ethics Board
under the application number
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
{MREB\#: 5219}.

@@

\begin{figure}[!ht]
\centering
\includegraphics[scale=0.5]{PrimaryLanguages.pdf}
\centering
\caption{\label{fig_language}Primary languages versus number of projects}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{installability_scores.pdf}
\caption{AHP installability scores}
\label{fg_installability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{correctness_verifiability_scores.pdf}
\caption{AHP correctness \& verifiability scores}
\label{fg_correctness_verifiability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{reliability_scores.pdf}
\caption{AHP surface reliability scores}
\label{fg_reliability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{robustness_scores.pdf}
\caption{AHP surface robustness scores}
\label{fg_robustness_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{usability_scores.pdf}
\caption{AHP surface usability scores}
\label{fg_usability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{maintainability_scores.pdf}
\caption{AHP maintainability scores}
\label{fg_maintainability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{reusability_scores.pdf}
\caption{AHP reusability scores}
\label{fg_reusability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{understandability_scores.pdf}
\caption{AHP surface understandability scores}
\label{fg_surface_understandability_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{visibility_transparency_scores.pdf}
\caption{AHP visibility/transparency scores}
\label{fg_visibility_transparency_scores}
\end{figure}

@@

\begin{figure}[!ht]
\includegraphics[scale=0.47]{overall_scores.pdf}
\caption{Overall AHP scores with an equal weighting for all 9 software qualities}

@@

\begin{table}[H]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \citet{USGS2019} & \citet{TobiasEtAl2018} & \citet{BrettEtAl2021} &
\citet{WilsonEtAl2016} & \citet{SmithAndRoscoe2018} & \citet{HerouxEtAl2008} &
\citet{ThielEtAl2020} & \citet{vanGompelEtAl2016} & \citet{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark &
\checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & &
\checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark &
\checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark &
\checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark & R\\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark &
\checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark &
\checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table}

@@
